---
title: "Tue workshop"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working with Range Data
==============================

# Set up 
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()
BiocManager::available()
biocLite("GenomicRanges")
```

#IRanges
```{r}
library(IRanges) #you might see some package startup messages when you run this

rng<- IRanges(start=4, end=13)
rng
```
#IRanges with specific width
```{r}
IRanges(start=4, width = 3)

IRanges(end=5, width=5)

```
#IRange constructor
```{r}
x<- IRanges(start=c(4, 7, 2, 20), end=c(13, 7, 5, 23))
x

names(x)<- letters[1:4] #name the range in letter
x
```
#Accessor functions
```{r}
start(x)
end(x)
width(x)
end(x)<- end(x)+4 
x
```
#Subset IRanges
```{r}
x[2:3] # The 2 and 3 row of x
x[start(x)<5] # This will show the start that smaller than 5
x[width(x)>8] # This will show the start that bigger than 8
x['a'] # The row with name "a"

a<- IRanges(start = 7, width=4)
b<- IRanges(start = 2, end = 5)
c(a,b) # Using the function c() to merge ranges
```

#Basic Range Operations:Arithmetic, Transformations, and Set Operations
```{r}
x<- IRanges(start=c(40, 80),end = c(67,114))
x+4L
x-10L

y<- IRanges(start = c(4, 6, 10, 12), width=13)
y
restrict(y, 5, 10) # cuts a set of ranges such that they fall inside of certain bound

flank(x, width = 7) #flank creates ranges width positions upstream of the ranges 
flank(x, width = 7, start=FALSE) #By setting start=false, it flanking ranges downstream

```
#Reduce

```{r}
set.seed(0)
alns<- IRanges(start = sample(seq_len(50),20), width = 5)
head(alns, 4) 
reduce(alns) # reduce takes a set of possibly overlapping ranges and reduces them to a set of nonoverlapping ranges
```
#Gaps
```{r}
gaps(alns) #find the gap between the ranges

```
#Difference, intersection, union
```{r}
a<- IRanges(start=4, end=13)
b<- IRanges(start=12, end=17)
intersect(a, b)
setdiff(a, b)
union(b, a)
union(a, b)

```

# FindOverlaps
```{r}
qry<- IRanges(start = c(1, 26, 19, 11, 21, 7), end = c(16, 30, 19, 15, 24, 8), names = letters[1:6])
sbj<- IRanges(start=c(1, 19, 10), end=c(5, 29, 16), names = letters[24:26])
qry
sbj
hts<- findOverlaps(qry, sbj) # find the overlap between the qry and sbj
hts

names(qry)[queryHits(hts)]
names(sbj)[subjectHits(hts)] # find the name of each query and subject range with an overlap

hts_within<- findOverlaps(qry, sbj, type = "within") #limit our overlap results to only include query ranges that fall entirely within subject ranges
hts_within

findOverlaps(qry, sbj, select = "first")
findOverlaps(qry, sbj, select = "last")
findOverlaps(qry, sbj, select = "arbitrary")


sbj_it<- IntervalTree(sbj) # Creating an Interval Tree object from an IRanges object 
sbj_it
class(sbj_it)

findOverlaps(qry, sbj_it)
```
#Extract information
```{r}
as.matrix(hts) # Hits objects can be coerced to matrix
countQueryHits(hts) # How many subject ranges each query IRanges object overlaps
setNames(countQueryHits(hts), names(qry))
countSubjectHits(hts) # How many query ranges overlap the subject ranges
setNames(countSubjectHits(hts), names(sbj))
ranges(hts, qry, sbj)
```
# SubsetByOverlaps and countOverlaps
```{r}
countOverlaps(qry, sbj) #countoverlaps is similar to the soultion using countQueryOverlaps and setnames
subsetByOverlaps(qry, sbj) #it is same as qry[unique(queryHits(hts))]
```
# Finding Nearest Ranges and Calculating Distance
```{r}
qry<- IRanges(start=6, end = 13, names = 'query')
sbj<- IRanges(start = c(2, 4, 18, 19), end = c(4, 5, 21, 24), names = 1:4)
qry
sbj
nearest(qry, sbj)
precede(qry, sbj)
follow(qry, sbj)
```





0-based coordinate system 

- "CTTACTTCGAAGGCTG" [1:5]

# 1-based coordinate system 
```{r}
substr("CTTACTTCGAAGGCTG", 2, 5)




